Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-05-26 paper/graph.tex
Added bibliography
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@619 3fc257bb-8550-0410-81b3-a5c103da83f7

A triangle in a graph is any triplet of vertices $(V_i,V_j,V_k)$ where
the edges $(V_i,V_j), (V_j,V_k), (V_i,V_k)$ exist.  Figure
\ref{fig:tri} outlines a MapReduce algorithm that enumerates all
triangles, assuming an input graph of undirected edges $(V_i,V_j)$
where $V_i < V_j$ for every edge, i.e. an upper-triangular R-MAT
matrix.  This exposition follows the triangle-finding algorithm
presented in \cite{Cohen}.

@@

A connected component of a graph is a set of vertices where all pairs
of vertices in the set are connected by a path of edges.  A sparse
graph may contain many such components.  Figure~\ref{fig:cc} outlines
a MapReduce algorithm that labels each vertex in a graph with a
component ID.  All vertices in the same component are labelled with
the same ID, which is the ID of a vertex in the component.  We assume
an input graph of undirected edges $(V_i,V_j)$.  This exposition also
follows the connected-component algorithm presented in \cite{Cohen},
with the addition of logic that load-balances data across processors
when one or a few giant components exist in the graph.

@@

An ``independent'' set of vertices in a graph is one where no pair of
vertices in the set share an edge.  The set is ``maximal'' if no
additional vertex can be added \footnote{A maximal set is ``maximum''
if it is the largest maximal set.  Finding a graph's maximum
independent set is an NP-hard problem, which MapReduce is unlikely to
help with.}.  Finding a maximal independent set (MIS) is useful in
several contexts, such as identifying large numbers of indepenent
starting vertices for graph traversals.  Luby's algorithm \cite{Luby}
is a well-known parallel method for finding a MIS.  Figure
\ref{fig:luby} outlines a MapReduce version of Luby's algorithm,
assuming an input graph of undirected edges $(V_i,V_j)$ where $V_i <
V_j$ for every edge, i.e. an upper-triangular R-MAT matrix.

@@

In MapReduce, global information about edge adjacencies and labeled vertices
is not available.  Instead, following the Bellman-Ford-style 
algorithm of~\cite{SSSPMapReduce, BellmanFord},
every edge of the graph is visited in each iteration,
although labels for an edge's vertices are updated only if the breadth-first
traversal has reached those vertices.  The algorithm begins with a {\it map()}
function loading edges $(V_i, V_j)$ and source vertex $V_s$ with distance 0.
In each iteration, the edges $(V_i, V_j)$ are collated with respect to
$V_i$, and a {\it reduce()} operation updates each $V_j$'s distance if
$V_i$'s distance is non-zero.  The algorithm is fully described in
Figure~\ref{fig:sssp}.

>>>>

A triangle in a graph is any triplet of vertices $(V_i,V_j,V_k)$ where
the edges $(V_i,V_j), (V_j,V_k), (V_i,V_k)$ exist.  Figure
\ref{fig:tri} outlines a MapReduce algorithm that enumerates all
triangles, assuming an input graph of undirected edges $(V_i,V_j)$
where $V_i < V_j$ for every edge, i.e. an upper-triangular R-MAT
matrix.  This exposition follows the triangle-finding algorithm
presented in~\cite{Cohen09}.

@@

A connected component of a graph is a set of vertices where all pairs
of vertices in the set are connected by a path of edges.  A sparse
graph may contain many such components.  Figure~\ref{fig:cc} outlines
a MapReduce algorithm that labels each vertex in a graph with a
component ID.  All vertices in the same component are labelled with
the same ID, which is the ID of a vertex in the component.  We assume
an input graph of undirected edges $(V_i,V_j)$.  This exposition also
follows the connected-component algorithm presented in~\cite{Cohen09},
with the addition of logic that load-balances data across processors
when one or a few giant components exist in the graph.

@@

An ``independent'' set of vertices in a graph is one where no pair of
vertices in the set share an edge.  The set is ``maximal'' if no
additional vertex can be added \footnote{A maximal set is ``maximum''
if it is the largest maximal set.  Finding a graph's maximum
independent set is an NP-hard problem, which MapReduce is unlikely to
help with.}.  Finding a maximal independent set (MIS) is useful in
several contexts, such as identifying large numbers of indepenent
starting vertices for graph traversals.  Luby's algorithm~\cite{Luby86}
is a well-known parallel method for finding a MIS.  Figure~\ref{fig:luby} 
outlines a MapReduce version of Luby's algorithm,
assuming an input graph of undirected edges $(V_i,V_j)$ where $V_i <
V_j$ for every edge, i.e. an upper-triangular R-MAT matrix.

@@

In MapReduce, global information about edge adjacencies and labeled vertices
is not available.  Instead, following the Bellman-Ford-style 
algorithm of~\cite{SSSPMapReduce, Bellman58, Ford62},
every edge of the graph is visited in each iteration,
although labels for an edge's vertices are updated only if the breadth-first
traversal has reached those vertices.  The algorithm begins with a {\it map()}
function loading edges $(V_i, V_j)$ and source vertex $V_s$ with distance 0.
In each iteration, the edges $(V_i, V_j)$ are collated with respect to
$V_i$, and a {\it reduce()} operation updates each $V_j$'s distance if
$V_i$'s distance is non-zero.  The algorithm is fully described in
Figure~\ref{fig:sssp}.

### 2010-05-26 paper/intro.tex
Added bibliography
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@619 3fc257bb-8550-0410-81b3-a5c103da83f7

MapReduce is the programming paradigm popularized by Google
researchers Dean and Ghemawat \cite{Dean}.  Their motivation was to
enable rapid development and deployment of analysis programs to
operate on massive data sets residing on Google's large distributed
clusters.  They introduced a novel way of thinking about certain kinds
of large-scale computations as a ``map'' operation followed by a
``reduce'' operation.  The power of the paradigm is that when cast in
this way, a nominally serial algorithm now becomes two highly parallel
operations working on data local to each processor, sandwiched around
an intermediate data-shuffling operation that requires inter-processor
communication.  The user need only write serial code for the
application-specific map and reduce functions; the parallel data
shuffle can be encapsulated in a library since its operation is
independent of the application.

The Google implementation of MapReduce is a C++ library with
communication between networked machines via remote procedure calls.
It allows for fault tolerance when large numbers of machines are used,
and can use disks as out-of-core memory to process petabyte-scale data
sets.  Tens of thousands of MapReduce programs have since been written
by Google researchers and are a significant part of the daily compute
tasks run by the company \cite{Dean2}.

Similarly, the open-source Hadoop implementation of MapReduce
\cite{Hadoop}, has become widely popular in the past few years for
parallel analysis of large-scale data sets at Yahoo and other
data-centric companies, as well as in university and laboratory
research groups, due to its free availability.  MapReduce programs in
Hadoop are typically written in Java, though it also supports use of
stand-alone map and reduce kernels, which can be written as shell
scripts or in other languages.

More recently, MapReduce formulations of traditional number-crunching
kinds of scientific computational tasks have been described, such as
post-processing analysis of simulation data \cite{Tu}, graph
algorithmics \cite{Cohen}, and linear algebra operations \cite{Fox}.
The paper by Tu et al \cite{Tu} was particularly insightful to us,
because it described how MapReduce could be implemented on top of the
ubiquitous distributed-memory message-passing interface (MPI), and how
the intermediate data-shuffle operation is conceptually identical to
the familiar MPI\_Alltoall operation.  Their implementation of
MapReduce was within a Python wrapper to simplify the writing of user
programs.  The paper motivated us to develop our own C++ library built
on top of MPI for use in graph analytics, which we initially released
as open-source software in mid-2009 \cite{MRMPI}.  We have since
worked to optimize several of the library's underlying algorithms and
to enable its operation in out-of-core mode on larger data sets.
These algorithmic improvements are described in this paper and are
part of the current downloadable version \cite{MRMPI}.

@@

\item {\it C++ library using MPI for inter-processor communication:}
The user writes a (typically) simple main program which runs on each
processor of a parallel machine, making calls to the MR-MPI library.
For map and reduce operations, the library calls back to user-provided
map() and reduce() functions.  The use of C++ allows precise control
over the memory and format of data allocated by each processor during
a MapReduce.  Library calls for performing a map, reduce, or data
shuffle, are synchronous, meaning all the processors participate and
finish the operation before proceeding.  Similarly, the use of MPI
within the library is the traditional mode of MPI\_Send and
MPI\_Recv calls between processor pairs using large aggregated
messages to improve bandwidth performance and reduce latency costs.  A
recent paper by \cite{Dongarra} also outlines the MapReduce formalism
from an MPI perspective, though they advocate a more asyncrhonous
approach, using one-way communication of small messages.

\item {Small, portable:} The entire MR-MPI library is a few thousand
lines of standard C++ code.  For parallel operation, the program is
linked with MPI, a standard message passing library available on all
distributed memory machines.  For serial operation, a dummy MPI
library (provided) can be substituted.  As a library, it can be
embedded in other codes \cite{Titan,TitanURL} to enable them to
perform MapReduces.

@@

\item {\it Flexible programmability:} An advantage of writing a
MapReduce program on top of MPI, is that the user program can invoke
MPI calls directly, if desired.  For example, one-line calls to
MPI\_Allreduce are often useful in determining the status of an
iterative graph algorithm, as described in Section \ref{sec:graph}.
The library interface also provides a user data pointer as an argument
passed back to all callback functions, so it is easy for the user
program to store ``state'' on each processor, accessible during the
map and reduce operations.  For example, various flags can be stored
that alter the operation of a map or reduce function, as can richer
data structures, that accumulate the results.

@@

The remainder of the paper is organized as follows.  The next two
sections \ref{sec:mr} and \ref{sec:outcore} describe how in-core and
out-of-core MapReduce primitives are formulated as MPI-based
operations in the MR-MPI library.  Section \ref{sec:graph} briefly
describes the formulation of several common graph algorithms as
MapReduce operations.  Section \ref{sec:results} gives performance
results for these algorithms running on a parallel cluster for graphs
%KDDKDD CHECK SIZES
ranging in size from 1 million to 1 trillion vertices or edges.  In
this section, we highlight the performance and complexity trade-offs
of a MapReduce approach versus other more special-purpose algorithms.
The latter generally perform better but are harder to implement
efficiently on distributed memory machines, due to the required
explicit management of parallelism, particularly for large out-of-core
data sets.  Section~\ref{sec:lessons} summarizes some lessons learned
from the implementation and use of our library.

>>>>

MapReduce is the programming paradigm popularized by Google
researchers Dean and Ghemawat~\cite{MapReduce2004}.  Their motivation was to
enable rapid development and deployment of analysis programs to
operate on massive data sets residing on Google's large distributed
clusters.  They introduced a novel way of thinking about certain kinds
of large-scale computations as a ``map'' operation followed by a
``reduce'' operation.  The power of the paradigm is that when cast in
this way, a nominally serial algorithm now becomes two highly parallel
operations working on data local to each processor, sandwiched around
an intermediate data-shuffling operation that requires inter-processor
communication.  The user need only write serial code for the
application-specific map and reduce functions; the parallel data
shuffle can be encapsulated in a library since its operation is
independent of the application.

The Google implementation of MapReduce is a C++ library with
communication between networked machines via remote procedure calls.
It allows for fault tolerance when large numbers of machines are used,
and can use disks as out-of-core memory to process petabyte-scale data
sets.  Tens of thousands of MapReduce programs have since been written
by Google researchers and are a significant part of the daily compute
tasks run by the company~\cite{MapReduce2008}.

Similarly, the open-source Hadoop implementation of MapReduce~\cite{Hadoop}, 
has become widely popular in the past few years for
parallel analysis of large-scale data sets at Yahoo and other
data-centric companies, as well as in university and laboratory
research groups, due to its free availability.  MapReduce programs in
Hadoop are typically written in Java, though it also supports use of
stand-alone map and reduce kernels, which can be written as shell
scripts or in other languages.

More recently, MapReduce formulations of traditional number-crunching
kinds of scientific computational tasks have been described, such as
post-processing analysis of simulation data~\cite{Tu08}, graph
algorithmics~\cite{Cohen09}, and linear algebra operations~\cite{Ekanayake09}.
The paper by Tu et al.~\cite{Tu08} was particularly insightful to us,
because it described how MapReduce could be implemented on top of the
ubiquitous distributed-memory message-passing interface (MPI), and how
the intermediate data-shuffle operation is conceptually identical to
the familiar MPI\_Alltoall operation.  Their implementation of
MapReduce was within a Python wrapper to simplify the writing of user
programs.  The paper motivated us to develop our own C++ library built
on top of MPI for use in graph analytics, which we initially released
as open-source software in mid-2009~\cite{MRMPI}.  We have since
worked to optimize several of the library's underlying algorithms and
to enable its operation in out-of-core mode on larger data sets.
These algorithmic improvements are described in this paper and are
part of the current downloadable version~\cite{MRMPI}.

@@

\item {\it C++ library using MPI for inter-processor communication:}
The user writes a (typically) simple main program which runs on each
processor of a parallel machine, making calls to the MR-MPI library.
For map and reduce operations, the library calls back to user-provided
map() and reduce() functions.  The use of C++ allows precise control
over the memory and format of data allocated by each processor during
a MapReduce.  Library calls for performing a map, reduce, or data
shuffle, are synchronous, meaning all the processors participate and
finish the operation before proceeding.  Similarly, the use of MPI
within the library is the traditional mode of MPI\_Send and
MPI\_Recv calls between processor pairs using large aggregated
messages to improve bandwidth performance and reduce latency costs.  A
recent paper~\cite{Hoefler09} also outlines the MapReduce formalism
from an MPI perspective, though they advocate a more asyncrhonous
approach, using one-way communication of small messages.

\item {Small, portable:} The entire MR-MPI library is a few thousand
lines of standard C++ code.  For parallel operation, the program is
linked with MPI, a standard message passing library available on all
distributed memory machines.  For serial operation, a dummy MPI
library (provided) can be substituted.  As a library, it can be
embedded in other codes~\cite{Titan} to enable them to
perform MapReduces.

@@

\item {\it Flexible programmability:} An advantage of writing a
MapReduce program on top of MPI, is that the user program can invoke
MPI calls directly, if desired.  For example, one-line calls to
MPI\_Allreduce are often useful in determining the status of an
iterative graph algorithm, as described in Section~\ref{sec:graph}.
The library interface also provides a user data pointer as an argument
passed back to all callback functions, so it is easy for the user
program to store ``state'' on each processor, accessible during the
map and reduce operations.  For example, various flags can be stored
that alter the operation of a map or reduce function, as can richer
data structures, that accumulate the results.

@@

The remainder of the paper is organized as follows.  The next two
sections~\ref{sec:mr} and~\ref{sec:outcore} describe how in-core and
out-of-core MapReduce primitives are formulated as MPI-based
operations in the MR-MPI library.  Section~\ref{sec:graph} briefly
describes the formulation of several common graph algorithms as
MapReduce operations.  Section~\ref{sec:results} gives performance
results for these algorithms running on a parallel cluster for graphs
%KDDKDD CHECK SIZES
ranging in size from 1 million to 2 billion vertices or edges.  In
this section, we highlight the performance and complexity trade-offs
of a MapReduce approach versus other more special-purpose algorithms.
The latter generally perform better but are harder to implement
efficiently on distributed memory machines, due to the required
explicit management of parallelism, particularly for large out-of-core
data sets.  Section~\ref{sec:lessons} summarizes some lessons learned
from the implementation and use of our library.

### 2010-05-26 paper/thanks.tex
Added bibliography
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@619 3fc257bb-8550-0410-81b3-a5c103da83f7

\section{Acknowledgements}

Sandia is a multiprogram laboratory operated by Sandia Corporation, a
Lockheed Martin Company, for the United States Department of
Energy's National Nuclear Security Administration under contract
DE-AC04-94AL85000.

Hadoop folks.

>>>>

\section{Acknowledgements}
\label{sec:thanks}

We thank the following individuals for their contributions to this
paper: Greg Bayer and Todd Plantenga (Sandia) for explaining Hadoop
concepts to us, and for the Hadoop implementations and timings of
Section~\ref{sec:results}; Jon Cohen (NSA) for fruitful discussions
about his MapReduce graph algorithms~\cite{Cohen09}; Brian Barrett
(Sandia) for the PBGL results of Section~\ref{sec:results}; Jon Berry
(Sandia) for the MTGL results of Section~\ref{sec:results}, and for
his overall support of this work and many useful discussions.

Sandia National Laboratories is a multi-program laboratory operated by
Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
company, for the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-AC04-94AL85000.
